Brazil

### 2023-03-27 paper.qmd
initialise milestione table for arrivals
extracting KATL single day data

abstract: |
  This document is a sample illustrating the Quarto `ieeetran` template.
  It includes the key elements of a scientific articles 
  (references, equations, figures, tables, code, cross references).
  The template enables the generation of IEEE-formatted article from a Jupypter notebook.
keywords: [template, demo]
reference-section-title: References
bibliography: bibliography.bib
note: Extra footnote.
# filters:
#    - latex-environment
environments: [mytable, mywidetable]
---

@@

Despite the introduction of an initial market-based mechanism, immediate action to curb fuel burn and CO2 emissions rests with improvements of operational efficiency.

@@

Different approaches exists to detect and describe aircraft flight phases, e.g. recent machine learning algorithm [@sun2017flightphase].
This paper implements a heuristic approach with a focus on the detection of arrival traffic at the study airports.

@@

>>>>

abstract: |
  This document is a sample illustrating the Quarto `ieeetran` template.
  It includes the key elements of a scientific articles 
  (references, equations, figures, tables, code, cross references).
  The template enables the generation of IEEE-formatted article from a Jupypter notebook.
keywords: [template, demo]
reference-section-title: References
bibliography: bibliography.bib
note: Extra footnote.

execute: 
  echo: false

# filters:
#    - latex-environment
environments: [mytable, mywidetable]
---

@@

Despite the introduction of an initial market-based mechanism, immediate action to curb fuel burn and CO2 emissions rests with improvements of operational efficiency.
<!-- verbatim -->
ICAO has introduced the trajectory-based operations concept, which promises operational and environmental benefits to aviation. 
where the flight trajectory of an aircraft is flown as close as possible to the userpreferred route, with as little disruptions as possible through collaborative decision-making mechanism. This includes reducing potential conflicts and resolving demand/capacity imbalances earlier and more efficiently. TBO can therefore bring significant operational and environmental benefits to aviation.  
<!-- end verbatim copy from ICAO paper -->

@@

Different approaches exists to detect and describe aircraft flight phases, e.g. recent machine learning algorithm [@sun2017flightphase].
This paper implements a heuristic approach with a focus on the detection of arrival traffic at the study airports.
Figure @fig-EGLL-arrivals shows the detected arrival flights for a single day.

```{r}
#| label: fig-EGLL-arrivals
#| fig-cap: Arrivals at London Heathrow (EGLL) on sample day
#| out-width: 45%     # 2 column half-size
knitr::include_graphics("./figures/EGLL-arrivals-single-day.png")
```

@@

```{r}
#| label: fig-KATL-arrivals
#| fig-cap: Arrivals at Atlanta Hartsfiled (KATL) on sample day
#| out-width: 45%     # 2 column half-size
knitr::include_graphics("./figures/KATL-traffic-single-day.png")
```

### 2023-03-27 x_bra-check.qmd
extracting KATL single day data

```{r}
rwys_sbgr <- tribble(
   ~ICAO  , ~REF , ~LAT      , ~LON      , ~NAME    , ~ELEV
  , "SBGR", "ARP", -23.435555, -46.473056, "ARP"    , 2461
  , "SBGR", "O9L", -23.434281, -46.483361, "09L/27R", 2444
  , "SBGR", "27R", -23.424894, -46.448619, "09L/27R", 2440
  , "SBGR", "09R", -23.438797, -46.487027, "09R/27L", 2450
  , "SBGR", "27L", -23.431189, -46.458855, "09R/27L", 2445
)

#write_csv(rwys_sbgr, "./data/rwys-SBGR.csv")
```

```{r}
sbgr_arp   <- rwys_sbgr |> filter(ICAO == "SBGR", REF == "ARP") |> select(LAT, LON)
sbgr_200NM <- trrrj::polygon_at_distance(c(sbgr_arp$LON, sbgr_arp$LAT), d = 205) |> 
  sf::st_as_sf() |> 
  rename(geometry = x) |> 
  mutate(NAME = "SBGR-200")
```

```{r}
sbgr_base_plot <- ggplot() +
  geom_sf(data = sbgr_200NM) +
  geom_point(data = sbgr_arp, aes(x = LON, y = LAT))
sbgr_base_plot
```

@@

```{r}
# load helper scripts to download and store OSN data
source("./R/osn_dumps.R")
source("./R/osn_read_tardump.R")
source("./R/osn_make_nice_names.R")
```
```{r}
#| message: false
pth_to_dumps <- "../__DATA/OSN-MondayDumps"
ptn <- "2022-05-09"

@@

```{r}
spatial_subset <- function(.payload, .outer, ...){
  # check if payload is already a sf object
  if(!is(.payload,"sf")){
    .payload <- .payload |> cast_latlon_to_pts(.drop_coord = FALSE)
  }
  # spatial subsetting
  trimmed_data <- .payload[.outer,]
}

@@

```{r}
sbgr_base_plot + geom_sf(data = trjs_pts)
```

>>>>

```{r}
rwys_sbgr <- tribble(
   ~ICAO  , ~REF , ~LAT      , ~LON      , ~NAME    , ~ELEV
  , "SBGR", "ARP", -23.435555, -46.473056, "ARP"    , 2461
  , "SBGR", "09L", -23.434281, -46.483361, "09L/27R", 2444
  , "SBGR", "27R", -23.424894, -46.448619, "09L/27R", 2440
  , "SBGR", "09R", -23.438797, -46.487027, "09R/27L", 2450
  , "SBGR", "27L", -23.431189, -46.458855, "09R/27L", 2445
)

# write_csv(rwys_sbgr, "./data/rwys-SBGR.csv")
```

```{r}
rwys_sbgr  <- read_csv("./data/rwys-SBGR.csv", show_col_types = FALSE)
sbgr_arp   <- rwys_sbgr |> filter(ICAO == "SBGR", REF == "ARP") |> select(LAT, LON)
sbgr_200NM <- trrrj::polygon_at_distance(c(sbgr_arp$LON, sbgr_arp$LAT), d = 205) |> 
  sf::st_as_sf() |> 
  rename(geometry = x) |> 
  mutate(NAME = "SBGR-200")

# cast sf
arp_pt    <- sbgr_arp  |> cast_latlon_to_pts(.drop_coord = FALSE) 
rwys_pts  <- rwys_sbgr |> filter(REF != "ARP") |> cast_latlon_to_pts(.drop_coord = FALSE)
rwys_ls   <- rwys_sbgr |> cast_rwy_ls()

sbgr_centerline_ls <- cast_rwy_ctr_line_ls(rwys_sbgr |> filter(REF != "ARP"))
sbgr_rwy_box       <- airport_threshold_box(rwys_sbgr |> filter(REF != "ARP"))
sbgr_box           <- airport_centerline_box(rwys_sbgr |> filter(REF != "ARP"), .ctrl_length = 15000)
```

```{r}
sbgr_base_plot <- ggplot() +
  geom_sf(data = sbgr_200NM) +
  geom_sf(data = sbgr_box) +
  geom_point(data = sbgr_arp, aes(x = LON, y = LAT))
sbgr_base_plot
```

@@

```{}
# load helper scripts to download and store OSN data
source("./R/osn_dumps.R")
source("./R/osn_read_tardump.R")
source("./R/osn_make_nice_names.R")
```
```{}
#| message: false
pth_to_dumps <- "../__DATA/OSN-MondayDumps"
ptn <- "2022-05-09"

@@

```{}
spatial_subset <- function(.payload, .outer, ...){
  # check if payload is already a sf object
  # if not, coerce to sf
  if(!is(.payload,"sf")){
    .payload <- .payload |> cast_latlon_to_pts(.drop_coord = FALSE)
  }
  # spatial subsetting on global scale is calc intensive for sf
  # check for bounding box of .outer
  bbox <- sf::st_bbox(.outer)

  # simple pre-filter based on LAT/LON ~ bounding box
  trimmed_data <- .payload |> 
    dplyr::filter(LAT >= bbox$ymin, LAT <= bbox$ymax) |> 
    dplyr::filter(LON >= bbox$xmin, LON <= bbox$xmax)

  # spatial subsetting
  trimmed_data <- trimmed_data[.outer,]
}

@@

```{r}
trjs_pts <- arrow::read_parquet("./data/bra-sbgr-20220509.parquet") |> 
  coerce_unixepoch_to_datetime() |> 
  cast_latlon_to_pts(.drop_coord = FALSE)

sbgr_base_plot + geom_sf(data = trjs_pts)
```

Check for coverage close to SBGR

```{r}
arp_sbgr <- rwys_sbgr |> filter(REF == "ARP") |> select(LAT, LON)
apt_tfc  <- trjs_pts |> 
  trrrj::filter_positions_at_range(
    c(arp_sbgr$LON, arp_sbgr$LAT), 0, 15, lon = LON, lat = LAT
    ) |> 
  filter(ALT_B < 4000)

ggplot() + 
  geom_sf(   data = sbgr_rwys_ls, col = "blue" ) +
  geom_sf(   data = sbgr_box, fill = NA) +
  geom_point(data = apt_tfc, aes(x = LON, y = LAT), size = 0.5)
```

Which ids do we see at SBGR?

```{r}
sbgr_uids <- apt_tfc[sbgr_box,] |> distinct(ICAO24, FLTID)
viz_check_vertical(apt_tfc |> filter(ICAO24 %in% sbgr_uids$ICAO24))
```

```{r}
sbgr_trjs <- trjs_pts |> 
  filter(ICAO24 %in% sbgr_uids$ICAO24) |> 
  identify_trajectory_legs() |> 
  fix_invalid_fltid() |> 
  make_uid("2022-05-09")

ggplot() + geom_sf(data = sbgr_trjs)
```

```{r}
picks <- trjs_pts |> filter(FLTID %in% c("TAM3201", "GLO1612","TAM3172", "DLH506"))
viz_check_DAO(picks |> filter(ALT_B < 8000), sbgr_rwys_ls |> sf::st_buffer(dist = 10000)) + geom_sf(data = sbgr_rwys_ls)

viz_check_vertical(picks)
```
